Using Mean-Squared-Error in Log-Transformed Domain to Evaluate Speckle Filters.

### 2012-09-10 old.article.tex
before submit to TGRS?

%\usepackage[caption=false]{caption}%for captions
\usepackage[caption=false,font=footnotesize]{subfig} %for subfigures

%opening
\title{ SLC SAR speckle filtering using Homoskedastic effects of Logarithmic Transfomation }

%\author{Thanh-Hai Le, Ian McLoughlin}
\author{Thanh-Hai~Le,
        Ian~McLoughlin,~\IEEEmembership{Senior Member,~IEEE}, 
	Brian~Quang-Huy~Nguyen,
	Ken-Yoong~Lee, 
	and Timo~Brestchneider % <-this % stops a space
\thanks{Thanh-Hai~Le, Brian~Quang-Huy~Nguyen and Ian~McLoughlin are with the School of Computer Engineering, 
Nanyang Technological University, Singapore,}% <-this % stops a space
\thanks{Ken-Yoong Lee and Timo Brestchneider are with EADS InnovationWorks Singapore.}% <-this % stops a space
\thanks{Manuscript received ?, 2010; revised ?.}}

\markboth{Transactions on Geoscience \& Remote Sensing,~Vol.~?, No.~?, April~2010}%
{ Le \MakeLowercase{\textit{et al.}}: SLC SAR speckle filtering using Homoskedastic effects of Logarithmic Transfomation }

@@

\begin{abstract}
This paper note the ability of Logarithmic Transformation to convert heteroskedasticity of SAR single look data to homoskedastic values. Under the assumption of homogeneity, the sampling distribution of variance in log-transformed domain is constant and independent to the underlying back scattering coefficient. Then the observable sample variance in the log-transformed domain can be used as a measure of spatial non-homogeneity, thus statistical test on log-transformed variance can be formed to detect heterogeneity in a window of analysis. Using clustering algorithms, a novel speckle filtering algorithm is proposed to partition a non-homogenous area,  detected by the statistical test, into homogenous areas with estimated back scattering coefficient values.
\end{abstract}

\begin{IEEEkeywords}
Synthetic aperture radar, stochastic-process, speckle-filtering, inverse-problems, homoskedasticity
\end{IEEEkeywords}

%\textbf{Keywords: } Synthetic Aperture Radar, speckle-filtering, homoskedasticity

@@

\section{Introduction}
%\section{Logarithmic Transformation of SAR Single look data}

The nature of SAR speckle is stochastics over homogenous areas and heteroskedastic heterogenously. Statistical models are used to describe how the underlying back-scattering coefficient ($\sigma$) affect the distribution of measured SAR data. Speckle filtering, by and large, are estimators attempting to the unknown coefficient from observable SAR signal $(A)$.

SAR speckle-filtering can be and has been positioned within the context of estimation theory\cite{Touzi_2002_TGRS}. The statistical framework attempts to estimate unknown statistical parameters from observed data. The stages consist of statistical modelling, estimator development and evaluating of the performance of estimators. Estimators' are evaluated based on the bias and variance properties of their estimates. The performance evaluation is usually based qualitatively on some real data and quantitatively through simulated experiments. Due to the stochastic nature of simulation process, results of repeated experiments are normally preferred.

For about 30 years, speckle filtering has been an active research area, with new methods being introduced steadily (cite here some latest). That is because: eventhough, the statistical model within individual resolution cell is well established, the applicability of its corresponding estimators is restricted to homogenous areas. Practical images, however, are heterogenous. Crucially it is this spatial variation that is of high interest. This fact gives rise to the question that seemed obvious: how to call an analysis area heterogenous and subsequently what to do in the case of heterogeneity.

Various statistical models for heterogenous areas have been proposed (see \cite{Touzi_2002_TGRS} for a detailed review). Unfortunately, while most of the models highlight the multiplcative nature of sub-pixel or homogenous original SAR data, in extending the model to heterogenous images, virtually none has noted that spatial varition also gives rise to heteroskedasticity phenomena. Heteroskedastic, as explained in section ???, is defined as the dependence of conditional expected variance of original SAR data on the conditional expectation of its mean or equivalently its underlying back-scaterring coefficient $(\sigma)$. It is believed that heteroskedasticity give rise to serious negative impacts on various stages of speckle-filtering.

In modelling, heteroskedasticity has direct consequences in the central question of homogeneity or heterogeneity. In normal images, the contrast or variance among neighboring pixels have been used to measure homogeneity. Unfortunately such techniques do not appear to be effective under the heteroskedastic condition of original SAR data. In SAR images, both of these measures are dependent on the underlying coefficient $(\sigma)$. This make the problem of estimating variance is equal to the problem of estimating the mean and is the same as the main problem: estimating $\sigma$. The fact give rise to the vicious circle in SAR speckle filtering.

Case in point can be illustrated in the recently published improved sigma filter\cite{Lee_TGRS_2009}. The technique determines outlying points as being too far away from the standard deviation. However, as is done in \cite{Lee_TGRS_2009}, to estimate standard deviation, an estimator of mean is required and used. It is interesting to note that the MMSE estimator used to estimate the mean in \cite{Lee_TGRS_2009}, itself alone, is a rather successful speckle-filter\cite{Lee_PAMI_1980}.

Heteroskedasticity also poses numerous challenges in designing an efficient estimator. Heteroskedasticity directly violates Gauss-Markov theorem's homoskedastic assumption. Thus it renders the efficiency of any naive Ordinary Least Square estimator in serious doubt \cite{Furno_1991_JStatCompSimul}. If the variance is known \textit{a priori}, it has been proven, cite Aitken, that a weighted mean estimator is the best linear unbiased estimator. Interestingly, as noted by Lopes \cite{Lopes_TGRS_1990}, most known common successful adaptive filters \cite{Lee_PAMI_1980} \cite{Kuan_1985_PAMI} \cite{Frost_PAMI_1982} does make use of weighted mean estimators. The caveat however is that in SAR speckle filtering, variance is not known a priori. And eventhough variance can be estimated from observable values. And as the vicious circle goes, estimating the variance is as good as estimating the underlying coefficient $\sigma$ itself.

Heteroskedastic also decrease the effectiveness of estimators by increasing the variance of even best possible estimations. This fact makes it tricky to evaluate the performance of different estimators, even if quantitative experiments results is reported. In fact several quantitative simulated experiments has been carried out to compare the performance of different estimators\cite{Lee_TGRS_2009}. However only results of single experiment were reported. Due to the stochastic nature of the simulated experiement process, more experiments should be done to report the effectiveness, ie. variance, of each estimators' results.

Last but certainly not least, is the bad impact of heteroskedastic on SAR image interpretation. Most of the task to be carried out in intepreting SAR images almost certainly involves target detection, target segmentation and/or target classification. All of these tasks requires a good similarity or discriminant functions. Foundation to all these is the need of a consistence sense of distance. Unfortunately, by definition heteroskedasticity provided inconsistent measures of distance. This inconsistent sense of distance coupled with the failure of the ordinary least square regression methods are believed to cause a large class of artificial neural networks as well as a number of other computational intelligence methods to underperform.

In this paper, we aim to neutralize these various impacts. The paper is organized as follows. Section ??? will be devoted to discussion of the heteroskedastic phenomena in the original per-pixel SAR amplitude and intensity model. Logarithmic transformation is extended to the model and the homoskedastic property of it is asserted. Section ??? will discuss how sampling distribution of variance in the log-transformed domain can be used to model various degrees of homogeneity. Section ??? will presents a novel speckle filtering algorithm developed from heterogenous inference made possible by statistical testing of sample variances. Section ??? will present evaluation of our estimator both quantitatively through simulated experiments and qualitatively on real-life images.

\section{SAR speckle statistic model}

Useful, robust estimators require un-assuming models. This section will describe our model for speckle filtering. The first part will explain the heteroskedastic nature of original amplitude and intensity SAR data. In extending the original model to log-transformed domain, care has been taken to rely solely on rigorous mathematical and statistical principles.

@@

SAR speckle phenomena is explained as the interference of many coherent but dephased back-scaterring components, each reflecting from different and distributed elementary scaterers \cite{Oliver_ProcIEEE_1963, Leith_ProcIEEE_1971}. These interference can be considered as a random walk on the 2D complex plane \cite{Goodman_JOptSocAm_76}.  The random nature of the process arises due to the unknown random location, height, distance and thus random phase of each elementary scaterer and its response.

Assuming the Central Limit Theorem is applicable \cite{Goodman_Springer_1975}, then the real part $A_r$ as well as the imaginary part $A_i$ of the observed SAR signal $A$ can be considered as random variables from uncorrelated Gaussian distributed stochastic processes with zero means and indentical variances $\sigma^2/2$  \cite{Lee_CRCPress_2009}. Their probability density function (pdf) is given as:

@@

It then can be proved that the measurable amplitude $A=\sqrt{A_r^2+A_i^2}$ is a random variable of Rayleigh distributed and consequently intensity $I=A^2=(A_r^2+A_i^2)$ is a random variable of negative exponential distributed random process.

\begin{IEEEeqnarray}{l l l}
pdf(A) &=& \frac{2A}{\sigma^2}e^{ \left( -\frac{A^2}{\sigma^2} \right) }\\
pdf(I) &=& \frac{1}{\sigma^2}e^{\left( -\frac{I}{\sigma^2} \right) }
\end{IEEEeqnarray}

From statistical perspective, the multiplicative nature of amplitude and intensity data can be explained as follows. Consider two fixed, independent to $\sigma$ unit distributions given below:

@@

It is then trivial to prove that amplitude and intensity is simply a scaled version of these unit variables, ie. $A= \sigma A_1 $ and $I= \sigma^2 I_1 $. These relationships evidently manifest a mutiplicative nature. In fact, this condition has long been noted, but from different perspectives, in various SAR models including multiplicative model (cite JS Lee) and product model \cite{Jakeman_1980_JPhysAMathGen}.

If spatial homogeneity is defined as imaging scenes having the same back-scattering coefficient $\sigma$, then over a homogenous area, the measured values can then be considered as samples coming from a single stochastic process. Consequently, the population expected mean and variance of the four distributions are given in table \ref{tbl:orginal_sar_avg_var}

\begin{table}[!h]
\caption{ Mean and Variance of Original SAR data are both related to the scale factor $\sigma$ }
\label{tbl:orginal_sar_avg_var}
\normalsize
\centering
%\begin{center}

\begin{tabular}{|l|l|}
\hline
Mean & Variance \\
\hline
$avg(A_1) = \frac{ \sqrt{\pi}}{2}$ & $var(A_1) = \frac{(4-\pi)}{4}$ \\
$avg(I_1) = 1$ & $var(I_1) = 1$ \\
$avg(A) = \frac{\sqrt{\pi}}{2} \cdot \sigma $ & $var(A) = \frac{(4-\pi)}{4} \cdot \sigma^2 $ \\
$avg(I) = \sigma^2 $ & $ var(I) = \sigma^4$ \\
\hline
\end{tabular}

@@

From the above analysis, it is evident that amplitude as well as intensity SAR data suffer from hetoroskedastic phenomena, which is defined as the dependence of conditional expected variance of SAR data on the conditional expectation of mean. In the context of speckle filtering, table \ref{tbl:orginal_sar_avg_var} indicates the vicious circle: estimating variance is equal to estimating the mean and is as good as main problem, ie. estimating the unknown parameter $\sigma$

The formulas above has long been noted. In fact while pioneering the estimation of the equivalent number of look (ENL) index, Lee et al has noted the ratio of expected standard deviation to mean is a constant in both cases (ie. $snr(A)=\sqrt{\frac{4}{\pi}-1}$ and $snr(I)=1$). Here, we just offer a different intepretation, which sets the stage for the discussion of logarithmic transformtion.

@@

We propose to consider base 2 logarithmic transformtion of SAR orignal random variables

>>>>

\usepackage[center]{caption}%for captions
\usepackage[caption=false,font=footnotesize]{subfig} %for subfigures

%opening
\title{ 
	Using Mean-Squared-Error in Log-Transformed Domain to Evaluate Speckle Filters.
}

%\author{Thanh-Hai Le, Ian McLoughlin}
\author{Thanh-Hai~Le,
        Ian~McLoughlin, 
	Quang-Huy~Nguyen,
	Chan-Hua~Vun%
	%Ken-Yoong~Lee, 
	%and Timo~Brestchneider % <-this % stops a space
\thanks{Thanh-Hai~Le, Quang-Huy~Nguyen and Chan-Hua~Vun are with School of Computer Engineering, 
Nanyang Technological University, Singapore. Ian~McLoughlin is with School of Information Science and Technology,
University of Science and Technology of China.
}% <-this % stops a space
\thanks{The authors wish to thank Dr. Ken-Yoong Lee and Dr. Timo Brestchneider of EADS Innovation-Works Singapore for 
	their initial discussions and for providing us the RADAR-SAT2 imagery used in this paper. }% <-this % stops a space
\thanks{Manuscript received ?, 2012; revised ?.}}

\markboth{Transactions on Geoscience \& Remote Sensing,~Vol.~?, No.~?, ?~2013}%
{ Le \MakeLowercase{\textit{et al.}}: Using Mean-Squared-Error in Log-Transformed Domain to Evaluate Speckle Filters }

@@

\begin{abstract}
Logarithmic transformation has the potential to convert 
	the multiplicative and heteroskedastic statistical distribution of SAR speckle noise into 
	a simpler additive and homoskedastic distribution model.
From this homoskedastic model, within the log-transformed domain, 
	a few consistent measures of distance emerge.
With these, subtractive differences become consistent and predictable. 
We show, through experiments and analysis, that 
	the familiar mean-squared-error (MSE) criteria can reliably serve 
	as a single unifying performance indicator to evaluate SAR speckle filters. 
In homogeneous areas, we establish a mathematical relationship that links the variance of 
	log-transformed SAR data with the canonical Equivalent-Number-of-Looks index.
For heterogeneous scenes, the standard residual-based bias evaluation, also in the log-transformed domain, 
	is shown to be equivalent to the ratio-based radiometric preservation criteria, frequently employed in the 
	original domain. 
Through experiments as well as analysis, we show that lower MSE suggests better feature detection and classification
	ability, which can be an important requirement for subsequent post-filtering steps in practical SAR data processing.
Combining bias and variance evaluation, we propose and describe the use of MSE evaluation in log-transformed domain to evaluate SAR speckle 
	filters.
\end{abstract}

\begin{IEEEkeywords}
Synthetic aperture radar, speckle-filtering, homoskedasticity, mean-squared-error
\end{IEEEkeywords}

@@

\section{Introduction}

The nature of SAR speckle is that
	it is stochastic even if the underlying radiometry is constant. 
And when there are spatial radiometric variation, not only its expected value changes, its heteroskedastic variance varies as well.
Nevertheless, statistical models have been developed to derive the underlying back-scattering coefficient ($\sigma$) from measured SAR data. 
As such, speckle filters are, by and large, estimators that attempt to determine this unknown coefficient from observable SAR data.

SAR speckle-filtering can be, and has been, positioned within the context of estimation theory\cite{Touzi_2002_TGRS}. 
The stages in this statistical framework consist of statistical modeling, estimator development and evaluation of the estimators' performance. 
%Estimators are typically evaluated based on the bias and variance properties of their estimates: 
%	with lower bias and / or lower variance, hence lower MSE indicates better accuracy.
Estimators are typically evaluated based on the bias and variance properties of their estimates:
  lower bias and/or lower variance would lead to lower MSE, which then indicates better accuracy.
Ideally, estimator evaluation should be based both qualitatively on some real data and quantitatively through simulated experiments. 
In addition, due to the stochastic nature of SAR-processing, and thus its simulating process, 
	statistical summaries of repeatedly-simulated experiments are normally preferred to single-run results.

Our survey of SAR speckle filter research, however, indicates 
	a different picture from the standard practice of the statistical framework, 
	specifically, in the stage of performance evaluation. 
Due to the multiplicative and heteroskedastic nature of speckle noise, 
	bias and variance evaluation may not be the most useful measures. 
Thus, the standard evaluation metric of mean-squared-error is not easily applicable,
	prompting alternative evaluation criteria to be proposed.
In fact, our survey of relevant literature fails to reveal 
	a single universally agreed quantitative metric for the performance of speckle filters.

Most commonly, each newly proposed speckle filter tends to be published along with its own methodology for evaluating its 
	performance. 
As such, many papers lack a comparative basis beyond simple visual qualitative comparison on a few image scenes. 
While such visual comparison is useful, as an evaluative methodology, regrettably it lacks scientific objectivity. 
Some papers do present quantitative measurements. 
However, due to the lack of a standardized performance criteria, 
	the evaluation metrics can change significantly from one paper to the next.

In this paper, we propose and validate the use of MSE in the homoskedastic log-transformed domain to evaluate speckle filters. 
The paper is structured as follows.
Section \ref{sec:lit_review} briefly surveys related work in the published literature, 
Section \ref{sec:log_transform} provides a brief discussion of the logarithmic transformation.
Section \ref{sec:eval_homo} then illustrates the use of variance (an MSE component) in the log-transformed domain to 
	evaluate different filters' speckle suppression power over homogeneous areas.
Section \ref{sec:eval_hetero} describes the use of the same MSE to evaluate the performance of speckle filters in heterogeneous areas.
Section \ref{sec:practical_conjecture} hypothesizes the use of residual MSE in the log-transformed domain 
	as a criteria for choosing the most suitable speckle filter for a given satellite-captured SAR image.
Finally, Section \ref{sec:discussion} discusses the findings and presents concluding remarks.

\section{Related Work in Literature}
\label{sec:lit_review}

It is customary to divide the performance evaluation of speckle filters into two distinct classes of 
	homogeneous and heterogeneous region evaluation.
Across homogeneous areas, speckle filters are expected to estimate with negligible radiometric bias.
As such, evaluating speckle filters over homogeneous areas has traditionally been focused on evaluating the variation of the estimators'
	output (a.k.a the speckle suppression power).
In contrast, the methodologies used to evaluate speckle filters over heterogeneous areas are much more complicated, 
	due in part to the following difficulties.
The first is that of determining an absolute ground-truth against which quantitative criteria can be measured.
The second challenge is to then define a quantifiable metric that allows the performance of different speckle filters 
	to be measured and compared.

In general, any metric to evaluate speckle filters should be relevant to the normal usage of such filters.
Furthermore, the application of any speckle filter in a SAR processing framework
   should enable an improvement in the measurement, 
	detection or classification of the underlying radiometric features.
As shown in the subsequent sections, by applying log-transformation, 
	this overall requirement of speckle filtering can be further broken down into two smaller requirements.
On the one hand, speckle filters should preserve the underlying radiometric signal (namely the radiometric 
	preservation requirement).
On the other hand, they should reduce the variation of the additive noise (i.e. speckle suppression power).
These requirements can be measured and evaluated by determining, also in the log-transformed domain, the bias 
	and variance error of the output.
As the MSE evaluation is a combination of bias and variance error, it is therefore capable of evaluating 
	the general requirements of speckle filtering.

For homogeneous scenes where the underlying radiometry is assumed to be constant, 
	the filtered results are considered, statistically speaking, to be samples of 
	a single but complex stochastic process.
From a logarithmic transformation perspective, the radiometric preservation and speckle suppression requirements 
	of speckle filters can be judged using the familiar bias and variance evaluation.

One metric that can be used to detect radiometric distortion is the ratio between the estimated and the 
	original value $r = X_{est} / X_{org}$ \cite{Oliver_2004_SciTech} \cite{Medeiros_2003_IJRS}.
A somewhat similar metric is used in \cite{Wang_2005_MIPR} as $r_w = (X_{est} - X_{org} )/ X_{org}$.
In the log-transformed domain, the equivalent criteria for evaluation could be performed by a simple subtraction.
In other words, it is clear that the bias evaluation in the log-transformed domain 
	can be used to evaluate the radiometric preservation requirement of speckle filters.

Specifically for homogeneous scenes, Shi et al. \cite{Shi_IGARSS_1994} found that 
	in the original domain the ``standard'' filters (boxcar, Lee\cite{Lee_PAMI_1980}, 
	Kuan \cite{Kuan_1985_PAMI}, Frost\cite{Frost_PAMI_1982}, 
	MAP\cite{Lopes_IGARSS_1990}) and their enhanced versions \cite{Lopes_TGRS_1990} can achieve negligible bias. 
In this paper, we will also show that all of these standard filters preserve
	not only the expected radiometric values
	but also a number of subtractive and consistent measures of distance exhibited in the log-transformed domain.

Several metrics have been developed to evaluate speckle suppression power.
The most common measure is the Equivalent Number of Looks (ENL) index 
$ENL=avg(I)^2/var(I)$
that was proposed by Lee \cite{Lee_1981_CGIP}.
Another very similar metric is the ratio of mean to standard deviation, $R=avg(I)/std(I)$ \cite{Gagnon_SPIEProc_1997} 
In Section \ref{sec:eval_homo}, it will be shown that, for homogeneous areas, ENL is mathematically related to variance in the 
	log-transformed domain.
Subsequently, we propose the use of log-variance to evaluate the noise suppression power of speckle filters, 
	which is the primary criteria used to evaluate speckle filters over homogeneous area.

Real-life and practical images, however, are not homogeneous.
Thus there are a number of associated difficulties in evaluating speckle filters over heterogeneous scenes.
The first difficulty in evaluating speckle filters for heterogeneous scenes is to select the basis for comparison. 
It is trivially easy to estimate the underlying radiometric coefficient if an area is known to be homogeneous.
However, without simulation or access to solid ground-truth, it is practically impossible to do so for 
	real-life images.
And hence, the need for speckle filters estimation is warranted.

Without ground truth, one way to evaluate radiometric preservation of filters is to compute the ratio image mentioned above. 
Under a multiplicative model, the ratio image is expected to comprise solely of
    the noise being removed (i.e. it should be completely random). 
Being random, this should display as little ``visible'' structure as possible. 
However, when displaying such images for visual evaluation, the ratios that are smaller than unity are much 
harder to distinguish than those that are bigger \cite{Medeiros_2003_IJRS}. 
We therefore propose to adopt a log-transformed domain perspective 
  where the multiplicative noise would then be transformed into a more familiar additive model.
More importantly, the ratio image would then become a simple subtraction image.
The evaluation methodology for such adoption would not change. 
Only that the residual image would become linear and additive.
And thus it would be more natural to be displayed and evaluated visually (see \ref{sec:schema_log_images}).

Another way to evaluate speckle filters is by comparing the feature preservation characteristics 
of the original noised data and those of the filtered image. 
When there is no ground-truth given, the feature are estimated in both the original noised and the filtered images.
Evaluation would then determine how closely related the two feature maps are. 
Various methods may be applied to extract features; examples of those that have been used are the 
	Hough Transform \cite{Medeiros_2003_IJRS}, Robert gradient edge detector \cite{Gagnon_SPIEProc_1997} 
	and edge map \cite{Frost_PAMI_1982}.
While significant effort has been spent on these evaluations, serious doubts remain concerning the precision 
of these methodologies.
This is because feature extraction algorithms are only approximations,
	whose accuracy is not only dependent upon the characteristics of the original image 
		but also heavily affected by 
		the inherent noise.
Unfortunately, speckle filters invariably alter the noise characteristics.
Thus, without a clear understanding of these dependencies and no absolute ground truth, 
	using feature extraction algorithms to evaluate speckle filters would leave serious questions on 
	how to interpret the results, especially its accuracy.

Since SAR statistical models are quite well understood, 
	simulation experiments with known ground-truth can be employed 
		to evaluate speckle filters.
In this paper, we also make use of a methodology similar to the ones discussed above with two important changes.
Firstly, our simulated experiments offer absolute ground-truth.
Secondly, our threshold-based feature extraction algorithm is extremely simple.
More importantly, the dependency between the performance of the algorithm and the level of noise is well established. 
This performance can be 
	conveniently visualized by plotting the Receiver Operating Characteristics Curve (ROC).   
It can also be objectively quantified by measuring the Area Under this Curve (AUC). 
As shown in section \ref{sec:eval_hetero}, this standard and normalized criteria allows a comparative evaluation of 
	feature preservation capabilities for different speckle filters.

%There are two main ways that ground truths are used in evaluating speckle filters.
%A simple method is to embed speckle noise into an existing optical image, and then use the filters to estimate noise-free imagary from this. 
%Normally the image is large, with a number of different features and test conducted in once. 
%The benefit of this method is that a wide range of features can be tested, which is probably closer to the real-life situation. 
%The main drawback is that, since the noise embedding process is stochastic in nature, reporting only a single experimental is probably not providing a very representative result. 
%It is of course, also dependent upon the nature of the test image and noise embedded process employed.
%
%Another method is to evaluate using a patterned, structured and repeated ground truth which may be artificial or real.
%Since the structure is repeated many times, the combined results become more representative. Also, the patterns can be pre-designed and lead to the possibility of repeatable evaluations between research groups.
%The drawback is that only a single type of target can be tested per image. These types may normally be point targets, edges and lines (all features that are currently used to evaluate speckle filters).

Even with known ground-truth, evaluating metrics need to be defined for quantitative measurements.
This is the second difficulty in evaluating speckle filters for heterogeneous scenes.
Under the conditions of heterogeneity, the standard speckle filters still introduce radiometric loss, 
	normally at local or regional levels.
In fact, the common consensus is that a powerful speckle suppression filter (for example the boxcar filter) 
	is likely to perform poorly in terms of preserving underlying radiometric differences 
	(e.g. causing excessive blurring).
Furthermore, we have not found many published articles combining the evaluation
  of these two seemingly contradicting requirements.

Overall, many different methods and metrics have been proposed to evaluate various aspects of speckle filters. 
However it is clearly advantageous to have a single metric that is able to judge if one filter is better than another. 
Wang \cite{Wang_2005_MIPR} proposed using fuzzy membership to weight opinions of an expert panel. 
Although this provides a potential solution, we consider it to be tedious in implementation, fuzzy in concept and 
	subjective in nature.

%Similar to the ENL index in homogeneous area, in establishing an evaluation metric, 
%	it is desirable that the criteria should also be scene-independent\cite{Shi_IGARSS_1994}.
%The most common measure of difference in image processing methods is to compute some substractive distances, either in intensity or in amplitude domain. 
%However since SAR noise is multiplicative in the original domain, these differences are not only dependent on the noise level but also on the amplitude level of the signal itself.
%Shi \cite{Shi_IGARSS_1994} define sharpness as the ratio between
%Shi define the mid-point position of an edge as the intensity value between the nominal values of the two regions, $M= { E(X_t) + E(X_b)}/{2}$
%	and the sharpness of an edge as the slope between the two nominal intensities value, $S=E(X_t)-E(X_b)$, 
%	where $E()$ denotes expectation operator.

Another approach is to apply a universal mean squared error criteria into the context of SAR data. 
However since SAR data is heteroskedastic, which violates the assumptions of the Gauss-Markov theorem, 
	the use of MSE is not straightforward. 
Thus Gagnon \cite{Gagnon_SPIEProc_1997} suggested the use of
  the ratio between the expected mean of the signal and the RMSE of the removed noise. 
The metric is argued to be similar in interpretation to the standard signal-to-noise ratio (SNR). 
Others have suggested the use of normalized MSE, which is essentially the ratio between MSE and the expected mean.

%Thus we propose the use of MSE in the homoskedastic log-transformed domain to evaluate eps. 
%This is similar to the investigation of residual image in log-transformed domain and also agrees with the log-transformed variance metric in homogeneous area (assuming that the filters have no problem preserving the mean). We also show experimentally that this MSE measure is inversely correlated with the AUC index mentioned earlier, suggesting that the lower the MSE index a filter can achieve, the better its feature preservation capability can be.

In this paper, homoskedastic property is shown to be available after logarithmic transformation of SAR data.
As the important Gauss-Markov theorem becomes applicable in this transformed domain, 
	the use of MSE is shown to be again relevant for evaluating statistical estimators (i.e. speckle filters).
Intuitively, each components of the MSE measure, 
	namely bias and variance evaluation, can be mapped into 
	the requirements of radiometric preservation and speckle suppression for speckle filters.
Before the details of our evaluation methodology is presented and discussed, 
	Section \ref{sec:log_transform} provides a brief discussion on logarithmic transformation for SAR images.

\section{Logarithmic Transformation}
\label{sec:log_transform}

In 
%the first portion of 
this section, the SAR speckle model in the original domain is first shown to be multiplicative and heteroskedastic.
The impacts of heteroskedasticity on speckle filtering is then discussed.
Logarithmic transformation is then shown to convert original speckle data into an homoskedastic model, 
	where the noise is not only additive, but also independent of the underlying radiometric signal.
From the consistent and additive noise in log-transformed domain,
    the consistent measures of dispersion ($ln(I)-ln(avg(I))$), contrast ($ln(I_1)-ln(I_2)$ and 
	variance ($var(ln(I))$) are described.

@@

The SAR speckle phenomena is often explained as the interference of many coherent but dephased back-scattering components,
	each reflecting from different and distributed elementary scatterers \cite{Oliver_ProcIEEE_1963, 
	Leith_ProcIEEE_1971}. This interference can be considered as a random walk on the 2D complex plane 
	\cite{Goodman_JOptSocAm_76}.  The random nature of the process arises due to the unknown random location, 
	height, distance and thus random phase of each elementary scaterer and its response.

Assuming the Central Limit Theorem is applicable \cite{Goodman_Springer_1975}, the real part $A_r$ as well as 
the imaginary part $A_i$ of the observed SAR signal $A$ can then be considered as random variables from 
uncorrelated Gaussian distributed stochastic processes with zero means and 
identical variances $\sigma^2/2$  \cite{Lee_CRCPress_2009}. Their probability density function (pdf) is given as:

@@

It can then be proved that the measurable amplitude $A=\sqrt{A_r^2+A_i^2}$ is a random variable of Rayleigh 
distribution and consequently the intensity $I=A^2=(A_r^2+A_i^2)$ is a random variable of negative exponential 
distributed random process.

\begin{IEEEeqnarray}{l l l}
pdf(A) &=& \frac{2A}{\sigma^2}e^{ \left( -{A^2}/{\sigma^2} \right) }\\
pdf(I) &=& \frac{1}{\sigma^2}e^{\left( -{I}/{\sigma^2} \right) }
\end{IEEEeqnarray}

From a statistical perspective, the multiplicative nature of amplitude and intensity data can be explained as follows. Consider two fixed, independent to $\sigma$ unit distributions given below:

@@

It is then trivial to prove that amplitude and intensity are simply scaled versions of these unit variables, 
ie. $A= \sigma A_1 $ and $I= \sigma^2 I_1 $. 
These relationships evidently manifest a multiplicative nature. 
In fact, this condition has long been noted, but from different perspectives, in various SAR models including 
the multiplicative model \cite{Lee_1981_CGIP} and product model \cite{Jakeman_1980_JPhysAMathGen}.

If spatial homogeneity is defined as imaging scenes having the same back-scattering coefficient $\sigma$, 
then over a homogenous area, the measured values can then be considered as samples coming from a single stochastic 
process. Consequently, the population expected mean and variance of the four distributions are given in Table \ref{tbl:orginal_sar_avg_var}.

\begin{table}[!h]
\caption{Both the mean and the variance of original SAR data are related to the scale factor $\sigma$.}
\label{tbl:orginal_sar_avg_var}
\normalsize
\centering
%\begin{center}

\begin{tabular}{|l|l|}
\hline
Mean & Variance \\
\hline
$avg(A_1) = { \sqrt{\pi}}/{2}$ & $var(A_1) = {(4-\pi)}/{4}$ \\
$avg(I_1) = 1$ & $var(I_1) = 1$ \\
$avg(A) = {\sigma \cdot \sqrt{\pi}}/{2} $ & $var(A) = \sigma^2 \cdot {(4-\pi)}/{4} $ \\
$avg(I) = \sigma^2 $ & $ var(I) = \sigma^4$ \\
\hline
\end{tabular}

@@

From the above analysis, it is evident that amplitude as well as intensity SAR data suffer from hetoroskedastic 
phenomena, which is defined as the dependence of conditional expected variance of SAR data on the conditional 
expectation of the mean. 
In the context of speckle filtering, Table \ref{tbl:orginal_sar_avg_var} indicates the vicious cycle: 
estimating variance is equal to estimating the mean and to estimating the unknown parameter $\sigma$, 
i.e. the main problem.

The formulas above have long been noted. In fact while pioneering the estimation of the equivalent number of looks 
(ENL) index, Lee et al commented that the ratio of expected standard deviation to mean is a constant in both 
cases (ie. $snr(A)=\sqrt{\frac{4}{\pi}-1}$ and $snr(I)=1$). Here, we just offer a different intepretation, 
which sets the stage for the discussion of logarithmic transformtion.

\subsection{The effects of Heteroskedasticity on Speckle Filters}

%For about 30 years, speckle filtering has been an active research area, with new methods being introduced steadily. 
%That is because: 
Even though the statistical model within individual resolution cells is well established, the applicability of its 
models is restricted to homogenous areas. 
Practical images, however, are heterogenous. Crucially it is this spatial variation that is of high interest. 
This fact gives rise to a question that seemed obvious: how to label an analysis area as being heterogenous and 
subsequently, what to do in the case of heterogeneity.

Various statistical models for heterogenous areas have been proposed (see \cite{Touzi_2002_TGRS} for a detailed 
review). Unfortunately, while most of the models highlight the multiplcative nature of sub-pixel or homogenous 
original SAR data, in extending the models to heterogenous images, virtually none have noted that spatial variation 
also gives rise to the heteroskedasticity phenomena. 
Heteroskedasticity, as explained in the previous section, is defined as the dependence of conditional expected 
variance of original SAR data on the conditional expectation of its mean, or equivalently its underlying 
back-scattering coefficient $(\sigma)$.
It is believed that this heteroskedasticity gives rise to serious negative impacts at various stages of 
speckle-filtering.

In modelling, heteroskedasticity has direct consequences to the central question of homogeneity or heterogeneity. 
In normal images, the contrast or variance among neighboring pixels is often used to measure homogeneity. 
Unfortunately such techniques do not appear to be effective under the heteroskedastic condition of original SAR data. 
In SAR images, both of these measures are dependent on the underlying coefficient $(\sigma)$. 
This makes the problem of estimating variance to be equal to the problem of estimating the mean (i.e. equivalent 
to estimating $\sigma$).
%The fact give rise to the vicious circle in SAR speckle filtering.

Heteroskedasticity also poses numerous challenges in designing and evaluating an efficient estimator. 
Heteroskedasticity directly violates Gauss-Markov theorem's homoskedastic assumption. 
Thus it renders the efficiency of any naive Ordinary Least Square estimator \cite{Furno_1991_JStatCompSimul}, 
together with the normal Mean Squared Error evaluation criteria in serious doubt. 
If the variance is known \textit{a priori}, it has been proven that a weighted mean estimator is the 
best linear unbiased estimator. 
Interestingly, as noted by Lopes \cite{Lopes_TGRS_1990}, most known common successful adaptive 
filters \cite{Lee_PAMI_1980} \cite{Kuan_1985_PAMI} \cite{Frost_PAMI_1982} does make use of weighted mean estimators. 
However, the caveat is that in SAR speckle filtering, variance is not known \textit{a priori}. 
And even though variance can be estimated from observable values, as the vicious circle goes, 
estimating the variance is as good as estimating the underlying coefficient $\sigma$ itself.

A case in point is the recently improved sigma filter\cite{Lee_TGRS_2009}. 
The technique determines outlying points as being too far away from the standard deviation. 
However, as is done in \cite{Lee_TGRS_2009}, to estimate standard deviation, an estimator of mean is required 
and used. It is interesting to note that the MMSE estimator used to estimate the mean 
in \cite{Lee_TGRS_2009}, itself alone, is a rather successful speckle-filter\cite{Lee_PAMI_1980}.

Last, but certainly not least is the bad impact of heteroskedastic on SAR image interpretation. 
Most of the task to be carried out in intepreting SAR images almost certainly involves target detection, 
target segmentation and/or target classification. 
Each of these tasks require good similarity or discriminant functions. 
Fundamental to these is the need of a consistent measures of distance. 
Unfortunately, by definition, heteroskedasticity leads to inconsistent measures of distance. 
This inconsistent measure of distance, coupled with the failure of the ordinary least square regression methods, are
 believed to cause a large class of artificial neural networks as well as a number of other computational 
 intelligence methods for SAR classification to underperform.

@@

\begin{table}[t]
\caption{ Mean and Variance of Log Transformed values:
  here the variance is no longer dependent on \sigma}
\label{tbl:sar_log_domain_avg_var}
\normalsize
\centering
%\begin{center}
